---
title: "solarModel"
author: "Beniamino Sartini"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
library(solarr)
library(ggplot2)
```

## Model's fit 

### Specification 

Let's start by creating a specification object
```{r}
spec <- solarModel_spec$new()
spec$set_mean.model(arOrder = 1, maOrder = 0)
spec$set_seasonal.mean(monthly.mean = FALSE)
spec$set_seasonal.variance(monthly.mean = FALSE)
spec$specification("Bologna", from="2005-01-01", to="2022-01-01")
```

```{r, echo=FALSE, results='asis'}
library(fansi)
# Run your colored output function, capturing the console text
txt <- capture.output(spec)
# Convert ANSI color codes (\[1;31m etc.) to HTML <span> styles
html_txt <- fansi::sgr_to_html(txt)
# Print as HTML so it renders properly
cat(html_txt, sep = "<br>\n")
```

### Estimation 

Then, a solarModel can be fitted
```{r}
# Iniaitlize a solarModel object
model <- solarModel$new(spec)
# Fit the model
#self <- model$.__enclos_env__$self
#private <- model$.__enclos_env__$private
model$fit()
```

```{r, echo=FALSE, results='asis'}
library(fansi)
# Run your colored output function, capturing the console text
txt <- capture.output(model)
# Convert ANSI color codes (\[1;31m etc.) to HTML <span> styles
html_txt <- fansi::sgr_to_html(txt)
# Print as HTML so it renders properly
cat(html_txt, sep = "<br>\n")
```
The estimated parameters are obtained 
```{r}
dplyr::bind_cols(model$coefficients) %>%
  tidyr::gather("parameter", "value") %>%
  knitr::kable()
```
and the standard errors
```{r}
bind_rows(model$spec$seasonal.mean$tidy, model$spec$mean.model$tidy, model$spec$seasonal.variance$tidy, model$spec$variance.model$tidy)
```

```{r}
model$spec$mixture.model$std.errors %>%
  knitr::kable()
```

### Active slots

```{r}
# Model's name 
model$spec$model_name
# Reference location name
model$spec$place
# Reference coordinates
model$spec$coords
```

Data 
```{r}
bind_rows(
  head(model$data, 5),
  tail(model$data, 5)
) %>%
  knitr::kable()
```

Seasonal data 
```{r}
bind_rows(
  head(model$seasonal_data, 5),
  tail(model$seasonal_data, 5)
) %>%
  knitr::kable()
```
Monthly data
```{r}
bind_rows(
  head(model$monthly_data, 5),
  tail(model$monthly_data, 5)
) %>%
  knitr::kable()
```
Conditional moments
```{r}
bind_rows(
  head(model$moments[[1]], 5),
  tail(model$moments[[1]], 5)
) %>%
  knitr::kable()
```

### QMLE estimation 

Total hessian and jacobians not implemented jet
```{r}
model$hessian
model$jacobian
```
With QMLE estimate the standard errors are updated inside each model
```{r}
model_QMLE <- solarModel_QMLE(model)
```

```{r}
bind_rows(model_QMLE$spec$seasonal.mean$tidy, 
          model_QMLE$spec$mean.model$tidy, 
          model_QMLE$spec$seasonal.variance$tidy, 
          model_QMLE$spec$variance.model$tidy)
```


## Diagnostic 

### Distribution Tests: Normality 

```{r}
solarModel_test_distribution(model, "n", type = "train") %>%
  knitr::kable()
```

### Distribution Tests: Gaussian Mixture 

```{r}
solarModel_test_distribution(model, "gm", type = "train") %>%
  knitr::kable()
```

### Autocorrelation Tests: Ljung-Box

```{r}
solarModel_test_autocorr(model, method = "lb", lag.max = 30) %>%
  knitr::kable()
```

### Autocorrelation Tests: Breush-Godfrey

```{r}
solarModel_test_autocorr(model, method = "bg", lag.max = 30) %>%
  knitr::kable()
```

### Autocorrelation Tests: Breush-Pagan

```{r}
solarModel_test_autocorr(model, method = "bp", lag.max = 30) %>%
  knitr::kable()
```

### PIT test 

```{r}
solarModel_PIT_test(model, ci = 0.05, type = "full") %>%
  knitr::kable()
```

### Log-predictive density 

```{r}

solarModel_test_LPD(model, "full") %>%
  knitr::kable()
```

### Forecasts tests

```{r}
solarModel_test_forecast(model, 0.9, "full") %>%
  knitr::kable()
```

## Scenarios 

```{r}
simSpec <- solarScenario_spec$new(model_QMLE, from = "2022-01-01", to = "2023-01-01", seed = 2, quiet = TRUE)
# Simulate scenarios 
scenario <- solarScenario$new(simSpec)
scenario$simulate_residuals(nsim = 10)
scenario$filter()
solarScenario_plot(scenario)
```

```{r}
df_emp <- scenario$spec$emp
df_emp$VaR <- solarScenario_VaR(scenario, 0.5)$data$VaR_alpha

mean(df_emp$GHI <= df_emp$VaR)
```


